Gravitational Lens Recovery with \Glass: Measuring the mass profile and shape of a lens

### 2014-01-29 limits/ms.tex
Updated inertia tensor to show full expression.
edits

\cite{1981ApJ...244..736Y} were remarkably prescient about the subsequent
development of lens modelling.  First, they introduced the technique of
choosing a parametric form for the lensing mass and then fitting for the
parameters, which is still the most common strategy \citep[see for
example][]{2010GReGr..42.2151K,2011A&ARv..19...47K}.  Second, they pointed out
the non-uniqueness of lens models -- lensing degeneracies.  Third, they
suggested combining lensing data with stellar kinematics and X-rays, to reduce
the effect of the degeneracies.  Later work, as well as following up these
suggestions, has introduced some further new ideas.  Four of these are
important for the present work:
%
\begin{enumerate}
\item {\bf Free-form modelling:} 
    In `free-form' or non-parametric modelling, there is no specified
    parametric form for the mass distribution.  There are still assumptions (or
    priors) on the mass distribution, such as smoothness or being centrally
    concentrated
    \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2009A&A...500..681M,2010ApJ...723.1678C}
    but these are much less restrictive than parametric forms
    (\cite{2006MNRAS.367.1209L} implemented a particularly elegant prior, which
    is that the mass distribution must be non-negative and no extra images are
    produced). To be concrete, we define from here on:

@@

  Being under-constrained, it is then {\it necessary} to explore model
  degeneracies rather than find a single `best-fit' solution. Free-form models
  are more commonly used with cluster lenses
  \citep{2006ApJ...652L...5S,2009ApJ...690..154S,2009A&A...500..681M,2013arXiv1304.2393S},
  but can be used with galaxy lenses as well, where their less restrictive
  assumptions can be important.  For example, in time-delay galaxy lenses,
  parametric models have historically led to tension with independent measures
  of the Hubble parameter $H_0$
  \citep[e.g.][]{2002astro.ph..4043K,2002ApJ...578...25K}; these are resolved
  once the less restrictive assumptions of free-form models are permitted
  \citep{2007ApJ...667..645R}.  \item Model ensembles, exploring a diverse
  range of possible mass distributions that nonetheless all fit the data,
  are a way of combating the non-uniqueness of models.  Such ensembles are
  possible in parametric models \citep{1999AJ....118...14B} but are more
  common in free-form models
  \citep{2000AJ....119..439W,2009ApJ...690..154S,2012MNRAS.425.3077L}.

\item {\bf Stellar kinematic constraints:} This was first suggested by
    \citet{2002MNRAS.337L...6T} as a means to break lensing degeneracies and
    determine cosmological parameters. This works because stellar kinematics
    can provide an independent estimate of the Einstein radius, via the virial
    theorem:
%
\begin{equation}
  \frac{\langle v^2_\mathrm{los}\rangle}{c^2} \approx
  \frac{\theta_E}{6\pi} \frac{D_S}{D_{LS}}
\label{eqn:virial}
\end{equation}
%
where $\langle v^2_\mathrm{los}\rangle$ is the projected stellar velocity
dispersion, and the above relation becomes exact for isothermal lenses.  The
use of two-dimensional kinematics by \cite{2011MNRAS.415.2215B} is especially
interesting.

@@

We can also more directly probe the recovery of the shape of the mass
distribution by defining a global measure of lens shape
%
\begin{equation} 
    s \equiv \lambda_1/\lambda_2
\end{equation} 
%
where $\lambda_1$ and $\lambda_2$ are the eigenvalues of the 2D inertia tensor
%
\begin{equation} 
    I_{ij} = (-1)^{1-\delta_{ij}}\sum_\theta M(\vec\theta) \theta_i\theta_j 
\label{eqn:inertia}
\end{equation}
%
and $\delta_{ij}$ is the Kronecker delta. We always take $\lambda_1$ to be the
largest value. As with $f_R$ and $f_\theta$, we only consider mass up to one
radial position passed the outermost image and compute the fractional error as:
%
\begin{equation} \label{ferror shape}
  f_\mathrm{shape} = \left|s - \tilde s\right|  / \tilde s
\end{equation}
%
where $\tilde s$ is the shape of the mock galaxy. The distribution of
$f_\mathrm{shape}$ for each mock galaxy and each test case is shown in
\figref{shape results}.  Interestingly, for this global shape parameter
recovery it appears more important to have time delay data and/or a central
image (T,M) than to have a quad or multiple sources with wide redshift
separation. In all cases, the stellar mass little-aids the recovery, reflecting
the fact that $s$ is heavily weighted towards the shape at the {\it edge} of
the mass map, rather than at the centre where the stars may dominate the
potential (see \eqnref{eqn:inertia}).

@@

%\cite{2007ApJ...667..645R} were the first to test the recovery of
%strong lensing surface density profiles using mock data generated from
%N-body models {\bf JR. Is this true?}. They found that when including
%time delays, circularly averaged surface density distributions can
%sufficiently well-recovered to make interesting comparisons with
%theoretical models Cosmological parameters, however, are more
%sensitive. Since then, several groups have used N-body mock data to
%explore the fidelity of strong lensing reconstructions
%\citep{2007MNRAS.380.1729L}. We expand on these earlier works
%by considering not only the recovery of the circularly averaged
%surface density profile of a lens, but also its shape, higher order
%measures of the mass recovery (a pixel-by-pixel comparison), and the
%utility of wrapping in constraints from stellar mass estimates and/or
%stellar kinematics.
%

>>>>

\cite{1981ApJ...244..736Y} were remarkably prescient about the subsequent
development of lens modelling.  First, they introduced the technique of
choosing a parametric form for the lensing mass and then fitting for the
parameters, which is still the most common strategy \citep[see for
example][]{2010GReGr..42.2151K,2011A&ARv..19...47K}.  Second, they pointed out
the non-uniqueness of lens models -- lensing degeneracies.  Third, they
suggested combining lensing data with stellar kinematics and X-rays, to reduce
the effect of the degeneracies.  Later work, as well as following up these
suggestions, has introduced some further new ideas.  Four of these are
important for the present work:
%
\begin{enumerate}
\item {\bf Free-form modelling:} In `free-form' or non-parametric
  modelling, there is no specified parametric form for the mass
  distribution.  There are still assumptions (or priors) on the mass
  distribution, such as smoothness or being centrally concentrated
  \citep{1997MNRAS.292..148S,2005MNRAS.360..477D,2009A&A...500..681M,2010ApJ...723.1678C}
  but these are much less restrictive than parametric forms.  A
  particularly elegant prior is implemented by
  \cite{2006MNRAS.367.1209L}, requiring that the mass distribution to
  be non-negative and no extra images allowed. To be concrete, we
  define from here on:

@@

  Being under-constrained, it is then {\it necessary} to explore model
  degeneracies rather than find a single `best-fit'
  solution. Free-form models are more commonly used with cluster
  lenses
  \citep{2006ApJ...652L...5S,2009ApJ...690..154S,2009A&A...500..681M,2013arXiv1304.2393S},
  but can be used with galaxy lenses as well, where their less
  restrictive assumptions can be important.  For example, in
  time-delay galaxy lenses, parametric models have historically led to
  tension with independent measures of the Hubble parameter $H_0$
  \citep[e.g.][]{2002astro.ph..4043K,2002ApJ...578...25K}; these are
  resolved once the less restrictive assumptions of free-form models
  are permitted \citep{2007ApJ...667..645R}.  Hybrid methods, using a
  mass grid on top a parametric model, have also been explored
  \citep[e.g.,][]{2010MNRAS.408.1969V}.

\item Model ensembles, exploring a diverse
  range of possible mass distributions that nonetheless all fit the data,
  are a way of combating the non-uniqueness of models.  Such ensembles are
  possible in parametric models \citep{1999AJ....118...14B} but are more
  common in free-form models
  \citep{2000AJ....119..439W,2009ApJ...690..154S,2012MNRAS.425.3077L}.

\item {\bf Stellar kinematic constraints:} This was first suggested by
  \citet{2002MNRAS.337L...6T} as a means to break lensing
  degeneracies. This works because stellar kinematics can provide an
  independent estimate of the Einstein radius, via the virial theorem:
%
\begin{equation}
  \frac{\langle v^2_\mathrm{los}\rangle}{c^2} \approx
  \frac{\theta_E}{6\pi} \frac{D_S}{D_{LS}}
\label{eqn:virial}
\end{equation}
%
where $\langle v^2_\mathrm{los}\rangle$ is the projected stellar
velocity dispersion, and the above relation becomes exact for
isothermal lenses.  The technique has been applied to many lenses
\citep{2006ApJ...649..599K,2008ApJ...682..964B}.  Going further, the
use of two-dimensional kinematics \citep{2011MNRAS.415.2215B} is
especially interesting.

@@

We can also more directly probe the recovery of the shape of the mass
distribution by defining a global measure of lens shape
%
\begin{equation} 
    s \equiv \lambda_1/\lambda_2
\end{equation} 
%
where $\lambda_1$ and $\lambda_2$ are the eigenvalues of the 2D inertia tensor
%
% \begin{gather}
%    I_{xx} = \sum_{\vec\theta} M(\vec\theta) \theta^2_y \quad ; \quad
%    I_{yy} = \sum_{\vec\theta} M(\vec\theta) \theta^2_x \\
%    I_{xy} = I_{yx} = - \sum_{\vec\theta} M(\vec\theta) \theta_x \theta_y \nonumber
%    \label{eqn:inertia}
% \end{gather}
%
\begin{equation}
\left(
\begin{matrix}
   \sum_{\vec\theta} M(\vec\theta) \theta^2_y 
    & - \sum_{\vec\theta} M(\vec\theta) \theta_x \theta_y \\
    - \sum_{\vec\theta} M(\vec\theta) \theta_x \theta_y
    & \sum_{\vec\theta} M(\vec\theta) \theta^2_x
\end{matrix}
\right).
    \label{eqn:inertia}
\end{equation}
%
We always take $\lambda_1$ to be the largest value. As with $f_R$ and
$f_\theta$, we only consider mass up to one radial position passed the
outermost image and compute the fractional error as:
%
\begin{equation} \label{ferror shape}
  f_\mathrm{shape} = \left|s - \tilde s\right|  / \tilde s
\end{equation}
%
where $\tilde s$ is the shape of the mock galaxy. The distribution of
$f_\mathrm{shape}$ for each mock galaxy and each test case is shown in
\figref{shape results}.  Interestingly, for this global shape parameter
recovery it appears more important to have time delay data and/or a central
image (T,M) than to have a quad or multiple sources with wide redshift
separation. In all cases, the stellar mass little-aids the recovery, reflecting
the fact that $s$ is heavily weighted towards the shape at the {\it edge} of
the mass map, rather than at the centre where the stars may dominate the
potential (see \eqnref{eqn:inertia}).

@@
